---
title: "analysis answer"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
```

Load the data...

```{r}
temperature <- read_table(
    "https://chryswoods.com/data_analysis_r/cetml1659on.txt",
    skip=6,
    na=c("-99.99", "-99.9"),
    col_types=cols("DATE"=col_integer())
)
```

Create the month levels

```{r}
month_levels <- c("JAN", "FEB", "MAR", "APR", "MAY", "JUN",
                  "JUL", "AUG", "SEP", "OCT", "NOV", "DEC")
```

Tidy the data...

```{r}
historical_temperature <- temperature %>%
    select(-YEAR) %>%
    pivot_longer(c("JAN", "FEB", "MAR", "APR", "MAY", "JUN",
                   "JUL", "AUG", "SEP", "OCT", "NOV", "DEC"),
                 names_to="month",
                 values_to="temperature") %>%
    rename(year=DATE) %>%
    mutate(month=factor(month, month_levels))
```

Calculate the mean monthly temperatures in the 17th Century

```{r}
c17th <- historical_temperature %>%
     filter(year<1700 & year>=1600) %>%
     group_by(month) %>%
     summarise("temperature"=mean(temperature, na.rm=TRUE), .groups="drop")
```

(the `.groups="drop"` removes a warning message in newer versions of R. It is
experimental, e.g. see [this stackoverflow post](https://stackoverflow.com/questions/62140483/how-to-interpret-dplyr-message-summarise-regrouping-output-by-x-override))

Calculate the mean monthly temperatures in the 21st Century

```{r}
c21st <- historical_temperature %>%
     filter(year>=2000) %>%
     group_by(month) %>%
     summarise("temperature"=mean(temperature, na.rm=TRUE), .groups="drop")
```

Now add the difference to the c21st table and print it out

```{r}
c21st["change"] <- c21st["temperature"] - c17th["temperature"]
c21st
```

From this we can see that most of the warming is focused on the winter
months.

We will now repeat this for the maximum and minimum temperatures...

```{r}
c17th_max <- historical_temperature %>%
     filter(year<1700 & year>=1600) %>%
     group_by(month) %>%
     summarise("temperature"=max(temperature, na.rm=TRUE), .groups="drop")
c21st_max <- historical_temperature %>%
     filter(year>=2000) %>%
     group_by(month) %>%
     summarise("temperature"=max(temperature, na.rm=TRUE), .groups="drop")
c21st_max["change"] <- c21st_max["temperature"] - c17th_max["temperature"]
c21st_max
```

```{r}
c17th_min <- historical_temperature %>%
     filter(year<1700 & year>=1600) %>%
     group_by(month) %>%
     summarise("temperature"=min(temperature, na.rm=TRUE), .groups="drop")
c21st_min <- historical_temperature %>%
     filter(year>=2000) %>%
     group_by(month) %>%
     summarise("temperature"=min(temperature, na.rm=TRUE), .groups="drop")
c21st_min["change"] <- c21st_min["temperature"] - c17th_min["temperature"]
c21st_min
```

Finally, we can get the average increase in monthly temperatures by calculating
the mean of the `change` column in `c21st`

```{r}
mean(c21st[["change"]])
```
