---
title: "files answer01"
output:
  html_document:
    df_print: paged
---

Import the tidyverse

```{r}
library(tidyverse)
```

Read in the file. As whitespace is the delimiter, we need to use `read_table`.
Note that `read_delim` with `delim=" "` is the wrong choice as it will try
to split on single whitespace characters. `read_table` is the right choice
for multiple whitespace separators.

Note that we should read the `DATE` as an integer, as it is a year.

```{r}
temperature <- read_table(
    "https://chryswoods.com/data_analysis_r/cetml1659on.txt",
    skip=6,
    na=c("-99.99", "-99.9"),
    col_types=cols("DATE"=col_integer())
)
temperature
```

How many years had a negative average temperature in January?

```{r}
negative_jan <- temperature %>% filter(temperature["JAN"] < 0)
negative_jan
```

```{r}
num_negative_jan <- as.numeric(count(negative_jan))
num_negative_jan
```

What was the average temperature in June over the years in the data set?

```{r}
jun_average <- mean(temperature[["JUN"]])
jun_average
```

(note that we have to get the column data, and not a 1-column tibble, 
hence why we use `[[ ]]`)
