---
title: "R Notebook"
output:   
  html_document:
    df_print: paged
---

First let's load and tidy the data, then add the "decade" variable.

```{r}
library(tidyverse)

temperature <- read_table(
    "https://chryswoods.com/data_analysis_r/cetml1659on.txt",
    skip=6,
    na=c("-99.99", "-99.9"),
    col_types=cols("DATE"=col_integer())
)

month_levels <- c("JAN", "FEB", "MAR", "APR", "MAY", "JUN",
                  "JUL", "AUG", "SEP", "OCT", "NOV", "DEC")

historical_temperature <- temperature %>%
    select(-YEAR) %>%
    pivot_longer(c("JAN", "FEB", "MAR", "APR", "MAY", "JUN",
                   "JUL", "AUG", "SEP", "OCT", "NOV", "DEC"),
                 names_to="month",
                 values_to="temperature") %>%
    rename(year=DATE) %>%
    mutate(month=factor(month, month_levels))

historical_temperature["decade"] <- (historical_temperature["year"] %/% 10) * 10
```

Here is the answer from the last section that has the "unpretty" graph...

```{r}
ggplot( historical_temperature %>%
          filter(month=="DEC") %>%
          group_by(decade) %>%
          summarise(average=mean(temperature, na.rm=TRUE), .groups="drop"),
        aes(x=decade, y=average)
      ) + 
  geom_point() +
  geom_smooth()
```

From [this chapter of the ggplot book](https://ggplot2-book.org/annotations.html)
and [here in the ggplot documentation](https://ggplot2.tidyverse.org/reference/labs.html)
we can see that we can add text annotations to the plot in many different ways.
The most useful for labelling the x and y axes is using 
[labs](https://ggplot2.tidyverse.org/reference/labs.html), [xlab](https://ggplot2.tidyverse.org/reference/labs.html) or
[ylab](https://ggplot2.tidyverse.org/reference/labs.html).

```{r}
ggplot( historical_temperature %>%
          filter(month=="DEC") %>%
          group_by(decade) %>%
          summarise(average=mean(temperature, na.rm=TRUE), .groups="drop"),
        aes(x=decade, y=average)
      ) + 
  geom_point() +
  geom_smooth() +
  labs(title="Average decadal temperature: Data originally from the Met Office",
       subtitle="(see https://chryswoods.com/data_analysis_r/files)") +
  xlab("Year") +
  ylab("Temperature / °C")
```

Note how we used the [unicode symbol for degrees](https://en.wikipedia.org/wiki/Degree_symbol)
to write `°C`. Text labels can be written in unicode, so can include emojis. 
Remember to be professional and restrain yourself from adding too many (or any?)
to production-quality graphs 😉.

Next, we can change the color of the points to red by changing the `colour` in 
its aesthetic.

```{r}
ggplot( historical_temperature %>%
          filter(month=="DEC") %>%
          group_by(decade) %>%
          summarise(average=mean(temperature, na.rm=TRUE), .groups="drop"),
        aes(x=decade, y=average)
      ) + 
  geom_point(color="red") +
  geom_smooth() +
  labs(title="Average decadal temperature: Data originally from the Met Office",
       subtitle="(see https://chryswoods.com/data_analysis_r/files)") +
  xlab("Year") +
  ylab("Temperature / °C")
```

From [this chapter of the ggplot book](https://ggplot2-book.org/scale-colour.html) we can 
see that ggplot supports assignment of colour scales to data. To start, we set the colour
aesthetic equal to a data series. In this case, we will colour the points representing
temperature by temperature. But you could equally colour by another variable in the 
data if that was appropriate.

```{r}
ggplot( historical_temperature %>%
          filter(month=="DEC") %>%
          group_by(decade) %>%
          summarise(average=mean(temperature, na.rm=TRUE), .groups="drop"),
        aes(x=decade, y=average, color=average)
      ) + 
  geom_point() +
  geom_smooth() +
  labs(title="Average decadal temperature: Data originally from the Met Office",
       subtitle="(see https://chryswoods.com/data_analysis_r/files)") +
  xlab("Year") +
  ylab("Temperature / °C")
```

(note that we had to put the colour aesthetic now in the ggplot function call
with the data)

A colour scale has been automatically added. It is not the best scale of colours.
We can see in the ggplot book that ggplot has a number of pre-defined colour
scales. The one we want is `RdBu`, which scales from red to blue.

To apply a colour scale we need to use either `scale_colour_brewer` (for discrete
data), or `scale_colour_distiller` (for continuous data). As the average
temperature is continuous, we use `scale_colour_distiller`.

```{r}
ggplot( historical_temperature %>%
          filter(month=="DEC") %>%
          group_by(decade) %>%
          summarise(average=mean(temperature, na.rm=TRUE), .groups="drop"),
        aes(x=decade, y=average, color=average)
      ) + 
  geom_point() +
  geom_smooth() +
  labs(title="Average decadal temperature: Data originally from the Met Office",
       subtitle="(see https://chryswoods.com/data_analysis_r/files)") +
  xlab("Year") +
  ylab("Temperature / °C") +
  scale_colour_distiller(palette = "RdBu")
```

The colour scale legend on the right is unnecessary, so we will remove it
using `theme(legend.position = "none")` (you can learn more about themes
in the [themes chapter of the ggplot book](https://ggplot2-book.org/polishing.html))

```{r}
ggplot( historical_temperature %>%
          filter(month=="DEC") %>%
          group_by(decade) %>%
          summarise(average=mean(temperature, na.rm=TRUE), .groups="drop"),
        aes(x=decade, y=average, color=average)
      ) + 
  geom_point() +
  geom_smooth() +
  theme(legend.position="none") +
  labs(title="Average decadal temperature: Data originally from the Met Office",
       subtitle="(see https://chryswoods.com/data_analysis_r/files)") +
  xlab("Year") +
  ylab("Temperature / °C") +
  scale_colour_distiller(palette = "RdBu")
```




