---
title: "tibbles answer 1"
output:
  html_document:
    df_print: paged
---

Import the tidyverse and load up the data

```{r}
library(tidyverse)

census <- tibble("City"=c("Paris", "Paris", "Paris", "Paris",
                          "London", "London", "London", "London",
                          "Rome", "Rome", "Rome", "Rome"),
                 "year"=as.integer(c(2001, 2008, 2009, 2010,
                                     2001, 2006, 2011, 2015,
                                     2001, 2006, 2009, 2012)),
                 "pop"=c(2.148, 2.211, 2.234, 2.244,
                         7.322, 7.657, 8.174, 8.615,
                         2.547, 2.627, 2.734, 2.627))
```

We start by grabbing the data for the year we care about

```{r}
census %>% filter(year==2001)
```

We can see that the smallest population was in Paris that year but 
let's try to extract it using R.

```{r}
pop <- (census %>% filter(year==2001))$pop
pop
```

The `min` function returns the minimum of a list of numbers. If we run
this on `pop` then we will get the smallest number.

```{r}
min_pop <- min(pop)
min_pop
```
We can now use this minimum population to further filter the census data;

```{r}
census %>% filter(year==2001) %>% filter(pop==min_pop)
```

Finally(!) we can extract the City column

```{r}
(census %>% filter(year==2001) %>% filter(pop==min_pop))["City"]
```
 
All of this could be combined into a single (dense) expression, e.g.

```{r}
city <- (census %>%
           filter(year==2001) %>%
           filter(pop==min((census %>% filter(year==2001))["pop"]))
         )["City"]
city
```

 
 