---
title: "Text and Sentiment Analysis in R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tokenising

```{r}
sherlock <- gutenberg_download(1661) %>% 
              mutate(linenumber=row_number())
```


Now create a table of all consequtive pairs of words - bigrams

```{r}
sherlock_bigrams <- sherlock %>% unnest_tokens(bigram, text, token = "ngrams", n = 2)
```

Now we want to remove pairs of words where one in the pair is a stop word. To do
this we must first separate the bigram column into two, which we will call
word1 and word2

```{r}
sherlock_bigrams <- sherlock_bigrams %>%      separate(bigram, c("word1", "word2"), sep = " ") 
```

```{r}
sherlock_bigrams <- sherlock_bigrams %>%      filter(!word1 %in% stop_words$word) %>%      filter(!word2 %in% stop_words$word)
```

Now let's also remove the NAs...

```{r}
sherlock_bigrams <- sherlock_bigrams %>%
                      filter(!is.na(word1)) %>% 
                      filter(!is.na(word2))
```

Finally(!) we will rejoin the word1 and word2 columns into a single "bigram" column, 
using the "unite" function

```{r}
sherlock_bigrams <- sherlock_bigrams %>%
                      unite(bigram, word1, word2, sep=" ")
```

Note that we could have written all of these steps as a single pipeline;

```{r}
sherlock_bigrams <- sherlock %>% 
                      unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
                      separate(bigram, c("word1", "word2"), sep = " ") %>%
                      filter(!word1 %in% stop_words$word) %>%                      filter(!word2 %in% stop_words$word) %>%
                      filter(!is.na(word1)) %>% 
                      filter(!is.na(word2)) %>% 
                      unite(bigram, word1, word2, sep=" ")
```

We can now count the bigrams as before, e.g.

```{r}
bigram_counts <- sherlock_bigrams %>% count(bigram, sort=TRUE)
```

This is good at finding names of important people and places, plus commonly used 
described items (e.g. "speckled band").

We could also use this to find specific streets by filtering on word2 (e.g. before we
join the bigrams together again, e.g.

```{r}
sherlock_streets <- sherlock %>%
                      unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
                      separate(bigram, c("word1", "word2"), sep = " ") %>%
                      filter(!word1 %in% stop_words$word) %>%                      filter(!word2 %in% stop_words$word) %>%
                      filter(!is.na(word1)) %>% 
                      filter(!is.na(word2)) %>% 
                      filter(str_detect(word2, "^st(reet)?$")) %>%
                      unite(bigram, word1, word2, sep=" ") %>%
                      count(bigram, sort=TRUE)
```

(see how the regular expression "^st(reet)?$" enabled us to filter both "st" and also
"street" in a single line - really they are worth learning ;-). You can build regexps
like this and test them using a regexp tester, like https://regex101.com).

## Visualising connections between words

```{r}
library(igraph)
```

```{r}
bigram_counts <- sherlock %>% 
                      unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
                      separate(bigram, c("word1", "word2"), sep = " ") %>%
                      filter(!word1 %in% stop_words$word) %>%                      filter(!word2 %in% stop_words$word) %>%
                      filter(!is.na(word1)) %>% 
                      filter(!is.na(word2)) %>% 
                      count(word1, word2, sort=TRUE)
```

Now create a graph

```{r}
bigram_graph <- bigram_counts %>%
                  filter(n >= 7) %>%
                  graph_from_data_frame()
```

Visualise using the ggraph package

```{r}
library(ggraph)

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1)
```

This isn't very pretty. By looking online at the options to improve
ggraph you can make it look better. The best is to copy a graph you like,
e.g. this one from here: https://www.tidytextmining.com/ngrams.html

```{r}
ggraph(bigram_graph, layout = "fr") +     geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,                    arrow = grid::arrow(type = "closed", length = unit(2, "mm")),                     end_cap = circle(1, "mm")) +     geom_node_point(color = "lightblue", size = 2) +     geom_node_text(aes(label = name), size = 2) +     theme_void()
```

This has surfaced extra information, e.g. we have found the Red Headed League!

Also, with this finer control we can reduce the font sizes etc enough that we can
look at more bigram associations, e.g.

```{r}
bigram_graph <- bigram_counts %>%    filter(n >= 3) %>%    graph_from_data_frame()

ggraph(bigram_graph, layout = "fr") +     geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,                    arrow = grid::arrow(type = "closed", length = unit(1, "mm")),                     end_cap = circle(0.5, "mm")) +     geom_node_point(color = "lightblue", size = 0.5) +     geom_node_text(aes(label = name), size = 1) +     theme_void()
```

Cool? :-)

Exercise

1. Create a similar bigram graph for the Hound of the Baskervilles (or another
book of your choice)

2. Create a bigram graph for the Hamlet Solil.. Don't remove the stop words, as
"to be" etc are very important, and also draw the graph of all bigrams
(don't filter based on count).

```{r}
lines <- readLines("https://chryswoods.com/text_analysis_r/hamlet.txt")hamlet <- tibble(line=1:length(lines), text=lines)

bigram_counts <- hamlet %>%     unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%    separate(bigram, c("word1", "word2"), sep = " ") %>%    filter(!is.na(word1)) %>%     filter(!is.na(word2)) %>%     count(word1, word2, sort=TRUE)

bigram_graph <- bigram_counts %>% 
                  graph_from_data_frame()

ggraph(bigram_graph, layout = "fr") +    geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,                   arrow = grid::arrow(type = "closed", length = unit(1, "mm")),                    end_cap = circle(0.5, "mm")) +    geom_node_point(color = "lightblue", size = 0.5) +    geom_node_text(aes(label = name), size = 1) +    theme_void()
```


## [Previous](wordclouds.html) | [Next](summary.html)

