---
title: "Text and Sentiment Analysis in R"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, error=FALSE)
```

# Regular Expressions

Searching for specific words or phrases from a block of text is relatively
easy. However, it can very complicated if you want to make this more flexible.
For example, imagine trying to find all of the titles and surnames
from this block of text;

```
Dear Mr. Johnson, 
  Dear Miss. Jameson,
Dear   Ms.   Jackson, 
Dear Mrs.    Peterson, 
  Dear    Mr. Sampson
Dear Dr.Johanson,
Dear Rev Richardson,
```

Searching and extracting text is a very complicated problem. Fortunately
it has been solved by computer scientists, and has been adopted
by most programming languages, including R. The solution is to use
what are called "regular expressions". 

## Regular Expressions in R

Regular expressions can look scary, but are pretty simple once you 
understand the rules. The syntax for regular expressions appeared 
and was standardised in the Perl language, and now nearly all 
programming languages support “Perl Compatible Regular Expressions” (PCRE).

In R, we use [str_detect](https://www.rdocumentation.org/packages/stringr/versions/1.4.0/topics/str_detect)
function from the [stringr](https://stringr.tidyverse.org) and
use [regex](https://stringr.tidyverse.org/articles/regular-expressions.html)
to pass in regular expressions. 

For example, let's first read Hamlet's soliloquy into a tibble.

```{r}
library(tidyverse)
lines <- readLines("https://chryswoods.com/text_analysis_r/hamlet.txt")
text <- tibble(line=1:length(lines), text=lines)
```

We can use `str_detect` to filter out lines that contain specific text,
e.g. here we will filter out all of the lines that contain the text `To`.

```{r}
text %>% filter(str_detect(text, "To"))
```

We can do the opposite of this, i.e. filter all of the lines that don't
contain `To` by negating the search.

```{r}
text %>% filter(str_detect(text, "To", negate=TRUE))
```

At the moment, this is performing a case-sensitive search. We can use
`regex` to build a regular expression that searches for `To` in 
a case-insensitive way.

```{r}
text %>% filter(str_detect(text, regex("To", ignore_case=TRUE)))
```

Regular expressions can do more than just search for specific text. They 
can search for patterns of text. For example, what if we wanted to search
for lines which contained the word `the`? We would want to exclude
lines where `the` is part of a word, like `their` or `whether`. To do this,
we use `\\b` to represent a word boundary (space, newline etc.). We can
then search for `\\bthe\\b` to search for `word boundary` followed by
`the` followed by `word boundary`.

```{r}
text %>% filter(str_detect(text, "\\bthe\\b"))
```

What about if we wanted to find all lines that contain `the` where `the` is
part of a word? In this case, we could use `\\w` to mean "any letter", and
thus `the\\w` would mean `the` followed by `any letter`.

```{r}
text %>% filter(str_detect(text, "the\\w"))
```

This has picked up `whether`, `rather` etc. What if we wanted all words that start
with `the`? We could combine `\\b` and `\\w` to search for lines that
contain `word boundary` followed by `the` followed by `any letter`, i.e.

```{r}
text %>% filter(str_detect(text, "\\bthe\\w"))
```

This has matched `them`, `there's` and `their`.

There are many special characters you can use, e.g.

* `\\d` Match any digit (number)
* `\\s` Match a space
* `\\w` Match any word character (alphanumeric and “_”)
* `\\S` Match any non-whitespace character
* `\\D` Match any non-digit character
* `.` Match any character
* `\\t` Match a tab
* `\\n` Match a newline

Instead of using `\\w` to mean any letter, we can actually ask to match
specific letters. We do this by putting the letters we want to match between
square brackets. For example, `[aiy]` would mean match any of `a`, `i` or `y`.
Thus `th[aiy]` means match `th` followed by `a`, `i` or `y`.

```{r}
text %>% filter(str_detect(text, "th[aiy]"))
```

You can control which characters are matched in the square brackets using;

* `[abc]` Match a, b or c
* `[a-z]` Match any character between a to z
* `[A-Z]` Match any character between A to Z
* `[a-zA-Z]` Match any character from a to z and A to Z (any letter)
* `[0-9]` Match any digit
* `[02468]` Match any even digit
* `[^0-9]` Matches NOT digits (^ means NOT)

You can also use repetition in your matching.

* `*` Match 0 or more times, e.g. `\\w*` means match 0 or more word characters
* `+` Match 1 or more times, e.g. `\\w+` means match 1 or more word characters
* `?` Match 0 or 1 times, e.g. `\\w?` means match 0 or 1 word characters
* `{n}` Match exactly `n` times, e.g. `\\w{3}` means match exactly 3 word characters
* `{n,}` Match at least `n` times, e.g. `\\w{5,}` means match at least 5 word characters
* `{m,n}` Match between `m` and `n` times, e.g. `\\w{5,7}` means match 5-7 word characters

We can use this to find all lines that contain words with 10-12 characters, by typing;

```{r}
text %>% filter(str_detect(text, "\\w{10,12}"))
```

You can also use this with sequences of characters, by putting them in 
round brackets. This is particularly useful when combined with `|`, which
means "or", e.g.

```{r}
text %>% filter(str_detect(text, "(monologue)|(dream)"))
```

has matched `monologue` or `dream`. You can combine this with the modifiers
above, e.g.

```{r}
text %>% filter(str_detect(text, "\\b(T|t)he(re)?\\b"))
```

matches a word boundary, followed by `T` or `t`, followed by `he` followed
by zero or one `re` followed by a word boundary. As such, it has matched
the words `The`, `the`, `There` and `there`.

You can also specify that a match must be made at the beginning of the 
line using `^`. This means match at the start of the line. Here we get
all of the lines that start with `The`.

```{r}
text %>% filter(str_detect(text, "^The"))
```

Similarly, use `$` to mean match at the end of the line. Here we get
all lines that end with `on`.

```{r}
text %>% filter(str_detect(text, "on$"))
```

Regular expressions are really powerful. There is great
[cheat sheet](https://github.com/rstudio/cheatsheets/blob/master/strings.pdf)
as part of stringr, and also a really nice description
[available here](https://stringr.tidyverse.org/articles/regular-expressions.html).

## Replacing text

As well as searching for text, you can also use regular expressions 
to replace text. You do this using stringrs 
[str_replace](https://stringr.tidyverse.org/reference/str_replace.html)
or [str_replace_all](https://stringr.tidyverse.org/reference/str_replace.html)
functions.

For example,

```
str_replace(line, pattern, replace)
```

will replace the first instance of `pattern` in `line` with `replace`, e.g.

```{r}
line <- text[4, 2]

str_replace(line, "be", "banana")
str_replace_all(line, "be", "banana")
```

You can also choose to match the string to replace using a regex, e.g.

```{r}
str_replace_all(line, regex("to be", ignore_case=TRUE), "ice-cream")
```

You need to use `mutate` to apply this change to all rows in a tibble.
For example;

```{r}
text %>% mutate(text = str_replace_all(text, "be", "banana"))
```

```{r}
text %>% mutate(text = str_replace_all(text, regex("\\bthe\\b", ignore_case=TRUE), "banana"))
```

## Health Warning

Regular expressions are very powerful. You can use them to search for specific 
output from your programs and to do powerful text manipulation. However, as you
have seen, they are very “write-only”. Extremely difficult to understand for
non-experts, and complex regular expressions can be difficult even for your
future-self to understand (i.e. “what was I thinking when I wrote that last
year? What does it mean and what does it do?”). You should ALWAYS comment
your regular expressions and explain in English exactly what you intended to
match when you wrote them. Once you have memorised the rules, you will find
regular expressions are very easy to read, use and are extremely powerful. 
However, without comments, they will be completely unintelligible to everyone 
else who looks at or relies on your code.

## Exercise

Find all words that follow `the` in Hamlet's soliloquy and replace them
with the word `banana`. 

[Answer](regexp_answer.html)

## [Previous](sentiment.html) | [Next](wordclouds.html)

