---
title: "Text and Sentiment Analysis in R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Regular Expressions

Take a little side-step into something that is really useful when analysing
and search text...

```{r}
lines <- readLines("https://chryswoods.com/text_analysis_r/hamlet.txt")
text <- tibble(line=1:length(lines), text=lines)
```

Use `str_detect` to look for particular lines

```{r}
text %>% filter(str_detect(text, "To"))
```

opposite search

```{r}
text %>% filter(str_detect(text, "To", negate=TRUE))
```

Use `regex` to give extra options, e.g. case insensitive match

```{r}
text %>% filter(str_detect(text, regex("To", ignore_case=TRUE)))
```

Lots of details about regexp here: https://stringr.tidyverse.org/articles/regular-expressions.html

```{r}
text %>% filter(str_detect(text, "\\bthe\\b"))
```

```{r}
text %>% filter(str_detect(text, "the\\w"))
```

```{r}
text %>% filter(str_detect(text, "\\bthe\\w"))
```

```{r}
text %>% filter(str_detect(text, "th[aiy]"))
```

```{r}
text %>% filter(str_detect(text, "\\w{10,12}"))
```

```{r}
text %>% filter(str_detect(text, "^The"))
```

```{r}
text %>% filter(str_detect(text, "on$"))
```

## Replacing text

Can use `str_replace` or `str_replace_all` to replace text in a line, e.g.

str_replace(line, pattern, replace)

will replace the first instance of 'pattern' in line with `replace`, e.g.

```{r}
line <- text[4, 2]

str_replace(line, "be", "banana")
str_replace_all(line, "be", "banana")
```

Can match using regex, e.g.

```{r}
str_replace_all(line, regex("to be", ignore_case=TRUE), "ice-cream")
```

Can use `mutate` to change the tibble, e.g.

```{r}
text %>% mutate(text = str_replace_all(text, "be", "banana"))
```

```{r}
text %>% mutate(text = str_replace_all(text, regex("\\bthe\\b", ignore_case=TRUE), "banana"))
```

## More info

Regex cheat sheet: https://github.com/rstudio/cheatsheets/blob/master/regex.pdf

(loads of R cheat sheets - https://github.com/rstudio/cheatsheets)

R regex tester app: https://adamspannbauer.github.io/2018/01/16/r-regex-tester-shiny-app/

General regex tester: https://regexr.com


## [Previous](sentiment.html) | [Next](wordclouds.html)

