---
title: "Text and Sentiment Analysis in R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, error=FALSE)
```

# Sentiment Analysis

Sentiment analysis involves trying to analyse the feelings or sentiments
expressed within a piece of text. Sentiments could be interpreted broadly,
e.g. with words being classed as "positive" or "negative", or more nuance
could be included, e.g. with words classed as expressing "joy", "fear",
"sadness" etc.

Sentiment analysis depends on building a dictionary of words and classifying
those words according to sentiment. Several such dictionaries have been 
created, typically via crowdsourcing or other large-scale analysis. The
[tidytext](https://www.rdocumentation.org/packages/tidytext/versions/0.2.6)
library provides the [get_sentiments](https://www.rdocumentation.org/packages/tidytext/versions/0.3.1/topics/get_sentiments) function. This can be used to download one of many
different sentiments dictionaries into a tibble.

Amongst others, tidytext provides access to three general purpose sentiments
dictionaries;

1. `AFINN` from [Finn Årup Nielsen](http://www2.imm.dtu.dk/pubdb/views/publication_details.php?id=6010),
2. `bing` from [Bing Liu and collaborators](https://www.cs.uic.edu/~liub/FBS/sentiment-analysis.html), and
3. `nrc` from [Saif Mohammad and Peter Turney](http://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm).

For example, we can download the `nrc` dictionary using;

```{r}
library(tidyverse)
library(tidytext)
sentiments <- get_sentiments("nrc")
sentiments
```

You can see that this really is just a dictionary of single words, with each
word assigned one or more sentiments (e.g. "sadness", "anger" etc.).

Note that this assignment of words to sentiments is based on individual
words only, and based on a fairly conservative modern usage of those words.
For example, "badly" is classified as both "negative" and "sadness", which, 
in my opinion, wouldn't capture the sentiment in sentences such as
"I badly want that!". Equally, "backwards" is classified as "disgust" and 
"negative", which is not how I would see it in sentences such as 
"they skillfully drove backwards around the course". So, when using these
dictionaries, be mindful of how they were created, the linguistic style
that they were generated against, and whether or not they would be 
appropriate for your text. You should also be careful of license, 
as different dictionaries are distributed with different licenses,
and different obligations on their use (e.g. citation, payment etc).

Note also that different dictionaries will use different sentiments,
and will classify them in different ways. For `nrc` the sentiments are
in the `sentiments` column. We can get the full list of sentiments and
the number of words classified to each one via;

```{r}
sentiments %>% count(sentiment)
```

## Matching sentiments to words

We can perform a sentiment analysis by attaching a sentiment to every
word in a block of text. For example, let's recreate the tidy text tibble
of all of the non-stopwords in the Advertures of Sherlock Holmes;

```{r}
library(gutenbergr)
data(stop_words)

sherlock <- gutenberg_download(1661) %>%
              unnest_tokens(word, text) %>%
              anti_join(stop_words)
sherlock
```

We can classify each word by sentiment by joining together the tidy text
`sherlock` tibble with the dictionary of sentiments in `sentiments`. We do
this using dplyr's [inner_join](https://dplyr.tidyverse.org/reference/join.html)
function. This will join together two tibbles via their common column 
(in this case, `word`), creating new rows as needed if a word appears twice.

```{r}
sherlock_sentiments <- sherlock %>% inner_join(sentiments)
sherlock_sentiments
```

Here you can see that this has added the `sentiment` column, and repeated
words where they have multiple sentiments, e.g. "scandal" being both
"fear" and "negative".

We can then count how many words of different sentiments there are
using [filter](https://dplyr.tidyverse.org/reference/filter.html)
and count.

```{r}
sherlock_sentiments %>% 
  filter(sentiment=="positive") %>%
  count(word, sort=TRUE)      
```

You could get the total number of words of each type using;

```{r}
sherlock_sentiments %>%
  filter(sentiment=="joy") %>%
  count(word) %>% 
  summarise(total=sum(n))
```

Alternatively (and much more simply) we could just count the number of 
occurrences of each sentiment in the sentiments column.

```{r}
sherlock_sentiments %>% count(sentiment)
```

We could plot this using a similar technique as the last section
(this time converting the sentiments into factors)

```{r}
sherlock_sentiments %>%
  count(sentiment) %>%
  mutate(sentiment = reorder(sentiment, n)) %>%
  ggplot(aes(n, sentiment)) + geom_col() + labs(y=NULL)
```

## Exercise

1. Perform a similar sentiment analysis for your chosen book from 
Project Gutenberg. How does the number of words of different sentiments
in your book compare to that for the Adventures of Sherlock Holmes.
Does this difference match what you would intuitively expect?

[Example answer here](sentiment_answer.html)

## Investigating change of sentiment through a text

You can also use sentiment analysis to investigate how sentiment change
throughout a text. To do this, we will first need to add a line number
to each word in our text. We can do this by mutating the tibble downloaded
from Project Gutenberg and adding a `linenumber` column which is just
the row number in the tibble (obtained via the 
[row_number](https://www.rdocumentation.org/packages/tidytable/versions/0.6.1/topics/row_number.) 
function).

```{r}
sherlock <- gutenberg_download(1661) %>% 
              mutate(linenumber=row_number()) %>% 
              unnest_tokens(word, text) %>% anti_join(stop_words)
sherlock
```

We can now add in the sentiments...

```{r}
sherlock_sentiments <- sherlock %>% 
                         inner_join(sentiments)
sherlock_sentiments
```

What we want to do now is to count the number of words with different sentiments
for differnet blocks of the text. One way to do this is to divide the text
up into blocks of, e.g 80 lines. We can count the number of sentiments
in each block, and then track how this count changes from block to block
in the text. Note that our choice of block size of 80 lines is arbitrary.
It is best to experiment with different block sizes to see what makes sense.

We can divide into block using `linenumber %/% 80`, which will modulo divide each
line number by 80. Thus lines 0-79 will be assigned to block 0, lines
80 to 159 to block 1 etc.

```{r}
sherlock_blocks <- sherlock_sentiments %>% 
                     count(block=linenumber%/%80, sentiment)
sherlock_blocks
```

Next, we can now pivot the tibble so that the sentiments are the 
columns, and we have one row per block of text. We do this using
the [pivot_wider](https://tidyr.tidyverse.org/articles/pivot.html)
function from the [tidyr](https://tidyr.tidyverse.org/index.html)
package. In this case, we want to pivot the tibble so that 
the names of the new columns come from the text values in
the current "sentiment" column, and the values in those columns
come from the current "n" column (which contains the number
of words of each sentiment). It may be that some sentiments
don't appear in a block, so we need to ensure that any missing
sentiments are filled in with a default value of 0.

```{r}
sherlock_blocks <- sherlock_blocks %>% 
                     pivot_wider(names_from = sentiment, 
                                 values_from = n, 
                                 values_fill = 0)
sherlock_blocks
```

We can now do things like calculate the difference between the number of 
positive and negative sentiments in each block, which could be placed
into a new column called "sentiment";

```{r}
sherlock_blocks <- sherlock_blocks %>%
                         mutate(sentiment = positive - negative)
sherlock_blocks
```

This could then be graphed, with the "sentiment" column on the y axis, 
and the block number on the x axis;

```{r}
sherlock_blocks %>% ggplot(aes(block, sentiment)) + geom_col()
```

This shows that the Adventures of Sherlock Holmes was more positive at
the beginning, but had more periods of negativity as the text progressed.

You can also look at the trajectory of specific sentiments, e.g.

```{r}
sherlock_blocks %>% ggplot(aes(block, joy)) + geom_col()
```

## Exercise

2. Draw a graph of the difference between positive and negative
   sentiments for your chosen book from Project Gutenberg.

[Example answer here](sentiment_answer.html)

3. Explore different number of lines within each block. How
   does this affect the graph?

[Example answer here](sentiment_answer.html)

## [Previous](tokenising.html) | [Next](regexp.html)

