---
title: "Text and Sentiment Analysis in R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tokenising

```{r}
install.packages("tidyverse")
install.packages("tidytext")
install.packages("textdata")
install.packages("gutenbergr")
```

```{r}
library(tidyverse)
```

```{r}
text = c("To be, or not to be--that is the question:",
         "Whether 'tis nobler in the mind to suffer",
         "The slings and arrows of outrageous fortune",
         "Or to take arms against a sea of troubles",
         "And by opposing end them.")
```

```{r}
text <- tibble(line=1:5, text=text)
```

```{r}
tokens <- text %>% unnest_tokens(word, text)
```

```{r}
tokens %>% count(word, sort=TRUE)
```

## Downloading texts

```{r}
library(gutenbergr)
```

```{r}
sherlock <- gutenberg_download(1661)
```

```{r}
sherlock %>% unnest_tokens(word, text) %>% count(word, sort=TRUE)
```

Dominated by "stop words"

```{r}
data(stop_words)
```

```{r}
sherlock %>% unnest_tokens(word, text) %>% anti_join(stop_words) %>% count(word, sort=TRUE)
```

(learn what an anti_join is)

```{r}
counts <- sherlock %>% 
            unnest_tokens(word, text) %>% 
            anti_join(stop_words) %>% 
            count(word, sort=TRUE) %>%
            filter(n > 75)
```

Need to convert words from character strings to factors if we want to plot them.
Use the `reorder` function.

```{r}
counts <- counts %>% mutate(word = reorder(word, n))
```

(word is now a factor, rather than a character string. This means that it can be used
as a category in a bar chart)

```{r}
counts %>% ggplot(aes(n, word)) + geom_col() + labs(y=NULL)
```

(aes puts n on the x axis, word on the y axis, used a bar/column chart, and then
dropped the label on the y axis)

## EXERCISE

Do the same analysis of the Hound of the Baskervilles and create the same graph.

Hound of the Baskervilles

```{r}
hound <- gutenberg_download(2852)
hound <- hound %>% 
           unnest_tokens(word, text) %>%
           anti_join(stop_words)
```

How does this visually compare to the Adventures of Sherlock Holmes?

## [Previous](index.html) | [Next](sentiment.html)

