---
title: "Text and Sentiment Analysis in R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Word clouds

Word clouds of word counts

```{r}
sherlock <- gutenberg_download(1661) %>%     mutate(linenumber=row_number()) %>%     unnest_tokens(word, text) %>% anti_join(stop_words)
```

```{r}
sherlock_counts <- sherlock %>% count(word, sort=TRUE)
```

```{r}
library(wordcloud)
wordcloud(counts$word, counts$n, min.freq=25, random.order=FALSE)
```

You can also use the `with` function to make this part of a pipeline, e.g.

```{r}
sherlock_counts %>% with(wordcloud(word, n, min.freq=25, random.order=FALSE))
```

You can also colour the words from highest frequency to least by adding
a colour palette, e.g.

```{r}
sherlock_counts %>% with(wordcloud(word, n, 
                                   min.freq=25, 
                                   random.order=FALSE,
                                   colours=brewer.pal(8, "Dark2")))
```

All of the available colour palettes are given here - http://applied-r.com/rcolorbrewer-palettes/

You can also display them all using

```{r}
display.brewer.all()
```

Exercise - draw a word cloud of your favourite book

## Colouring different sentiments

What about colouring words according to their sentiment. We would need to create the table
again that has the sentiment counts.

```{r}
sentiments <- get_sentiments("nrc")
sherlock_sentiments <- sherlock %>% 
                         inner_join(sentiments) %>% 
                         count(word, sentiment, sort=TRUE) %>% 
                         pivot_wider(names_from=sentiment, 
                                     values_from=n, 
                                     values_fill=0)
```

We could now draw word clouds for specific sentiments, e.g.

```{r}
sherlock_sentiments %>% with(wordcloud(word, joy, 
                                       min.freq=10,
                                       random.color = FALSE,
                                       colors = brewer.pal(8, "PuOr")))
```

```{r}
sherlock_sentiments %>% with(wordcloud(word, fear, 
                                       min.freq=10,
                                       random.color = FALSE,
                                       colors = brewer.pal(12, "Paired")))
```

Exercise

Create some sentiment word clouds for your chosen book, exploring
different sentiments. What are the advantages and disadvantages of word clouds?

## [Previous](sentiment.html) | [Next](ngrams.html)

