---
title: "Text and Sentiment Analysis in R"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, error=FALSE)
```

# Word clouds

Word clouds can be a useful visualisation to quickly view the most used
words in a block of text. R has the [wordcloud](https://www.rdocumentation.org/packages/wordcloud/versions/2.6/topics/wordcloud)
function in the wordcloud package that can be used to quickly generate
these plots.

To use it, we will first load the packages that will be used in this
chapter;

```{r}
library(tidyverse)
library(tidytext)
library(gutenbergr)
library(wordcloud)
```

We are going to create word clouds from the tidied text of
The Adventure of Sherlock Holmes. We start by downloading the text
and tidying it up, as we did in the previous chapters.

```{r}
sherlock <- gutenberg_download(1661) %>% 
    mutate(linenumber=row_number()) %>% 
    unnest_tokens(word, text) %>% anti_join(stop_words)
```

We want the size each word in the wordcloud to depend on the number
of times it appears. We thus create a tibble of these counts.

```{r}
sherlock_counts <- sherlock %>% count(word, sort=TRUE)
```

The wordcloud function is very easy to use. You just need to pass in a 
column of words (`sherlock_counts$word`) and a column of their
counts (`sherlock_counts$n`). We need to restrict the plot to only
displaying the common words, or else we will end up overloading R. We
do this by only displaying words that appear 25 or more times, using
`min.freq=25`. The `random.order` argument is set to `FALSE` so that
the plot is reproducible (as much as possible - the option tells R
to plot the words in order of descreasing frequency, rather than
a random order).

```{r}
wordcloud(sherlock_counts$word, sherlock_counts$n, 
          min.freq=25, random.order=FALSE)
```

You can also use the `with` function to make this part of a pipeline, e.g.

```{r}
sherlock_counts %>% with(wordcloud(word, n, 
                                   min.freq=25, random.order=FALSE))
```

You can also colour the words from highest frequency to least by adding
a colour palette, e.g.

```{r}
sherlock_counts %>% with(wordcloud(word, n, 
                                   min.freq=25, 
                                   random.order=FALSE,
                                   colors=brewer.pal(8, "Dark2")))
```

All of the available colour palettes are listed here - http://applied-r.com/rcolorbrewer-palettes/

You can also display the ones available in your R installation using;

```{r}
display.brewer.all()
```

## Exercise

Draw a coloured word cloud of words that appear in your chosen book.

## Colouring different sentiments

What about colouring words according to their sentiment? 
First we would need to recreate the table that has the sentiment counts,
using the same procedure as the previous chapter.

```{r}
sentiments <- get_sentiments("nrc")
sherlock_sentiments <- sherlock %>% 
                         inner_join(sentiments) %>% 
                         count(word, sentiment, sort=TRUE) %>% 
                         pivot_wider(names_from=sentiment, 
                                     values_from=n, 
                                     values_fill=0)
```

We could now draw word clouds for specific sentiments by picking out
specific columns, e.g. here are all of the common words associated with "joy".

```{r}
sherlock_sentiments %>% with(wordcloud(word, joy, 
                                       min.freq=10,
                                       random.color = FALSE,
                                       colors = brewer.pal(8, "PuOr")))
```

And here are the common words associated with "fear".

```{r}
sherlock_sentiments %>% with(wordcloud(word, fear, 
                                       min.freq=10,
                                       random.color = FALSE,
                                       colors = brewer.pal(12, "Paired")))
```

## Exercise

Create some sentiment word clouds for your chosen book, exploring
different sentiments. What are the advantages and disadvantages of word clouds?

## [Previous](regexp.html) | [Next](ngrams.html)
